Query Optimization for Dynamic Imputation

### 2017-07-13 paper/algorithm.tex
Replace Dirty(q) with C in imputation algs

The chained-equation algorithm proceeds by
iteratively fitting decision trees for the target attributes. 
In each iteration, the missing values of a single attribute are replaced with newly imputed values.  Values in a single column can be updated over multiple
iterations of the algorithm. With more iterations, the quality of the
imputation improves as values progressively reflect more accurate relationships
between attributes. The algorithm terminates after a fixed number of
iterations, or earlier if convergence is achieved, indicated by unchanged values in a 
column in subsequent iterations. For our experiments, we use Weka's \texttt{REPTree}
decision tree implementation~\cite{witten2016data}, a variant of the canonical \texttt{C4.5}
algorithm~\cite{quinlan1993}. Roughly, the time complexity of building one
decision tree is $\mathcal{O}(nm \text{ } log(n))$, where $n$ is the number of tuples
and $m$ is the number of dependent attributes~\cite{witten2016data}. Since we build
$k * |\textsc{Dirty}(q)|$ trees, we have
\begin{equation*}
    \begin{aligned}
    T(q) = k * &|\textsc{Dirty}(q)| * \textsc{Card}(q) * \\
               &(|\textsc{Clean}(q)| + |\textsc{Dirty}(q)| - 1) * log(\textsc{Card}(q)).
    \end{aligned}
\end{equation*}

Separately, we specify a heuristic penalty that takes into account the
improvement in imputation quality from more tuples and attributes, as well as
the diminishing returns from more data.
\[
    P(q) = \frac{|\textsc{Dirty}(q)|}{\sqrt{ \textsc{Attr}(q) * \textsc{Card}(q)}}
\]

@@

We can take advantage of the histograms maintained by \ProjectName{} to estimate the mean and construct a non-blocking impute operator.
Applying the heuristics from~\Cref{sec:quality} and~\Cref{sec:query-runtime}, we
have
\begin{equation*}
\begin{aligned}
    P(q) &= \sum_{c \in \textsc{Dirty}(q)} \frac{1+\widehat{Var}(c)}{\textsc{Card}(q)} \\
    T(q) &\approx 0.
\end{aligned}
\end{equation*}

@@

Random hot deck imputation is a non-parametric method in which missing values in a column are
replaced with randomly selected complete values from the same column, preserving the
distribution of the complete values. We use the same $P$ as for mean value imputation, but
we modify $T$ to account for the fact that hot-deck is blocking in our implementation.  
\begin{equation*}
\begin{aligned}
    P(q) &= \sum_{c \in \textsc{Dirty}(q)} \frac{1+\widehat{Var}(c)}{\textsc{Card}(q)} \\
    T(q) &= \textsc{Card}(q)
\end{aligned}
\end{equation*}
}

>>>>

The chained-equation algorithm proceeds by
iteratively fitting decision trees for the target attributes. 
In each iteration, the missing values of a single attribute are replaced with newly imputed
values. \cameraready{
Attributes that contain missing values but that are not being imputed are
ignored by the imputation.
}
Values in a single column can be updated over multiple
iterations of the algorithm. With more iterations, the quality of the
imputation improves as values progressively reflect more accurate relationships
between attributes. The algorithm terminates after a fixed number of
iterations, or earlier if convergence is achieved, indicated by unchanged values in a 
column in subsequent iterations. For our experiments, we use Weka's \texttt{REPTree}
decision tree implementation~\cite{witten2016data}, a variant of the canonical \texttt{C4.5}
algorithm~\cite{quinlan1993}. Roughly, the time complexity of building one
decision tree is $\mathcal{O}(nm \text{ } log(n))$, where $n$ is the number of tuples
and $m$ is the number of dependent attributes~\cite{witten2016data}.
\cameraready{
Recall that $C$ is the set of dirty attributes that are being imputed in a
given operation, where $C \subseteq \textsc{Dirty}(q)$.
With $k$ cycles of the algorithm, where in each cycle there are
$|C|$ iterations of fitting models for each dirty attribute, we have
}
\begin{equation*}
    \begin{aligned}
    T(q) = k * &|C| * \textsc{Card}(q) * \\
               &(|\textsc{Clean}(q)| + |C| - 1) * log(\textsc{Card}(q)).
    \end{aligned}
\end{equation*}

Separately, we specify a heuristic penalty that takes into account the
improvement in imputation quality from more tuples and attributes, as well as
the diminishing returns from more data.
\[
    P(q) = \frac{|C|}{\sqrt{ \textsc{Attr}(q) * \textsc{Card}(q)}}
\]

@@

We can take advantage of the histograms maintained by \ProjectName{} to estimate the mean and construct a non-blocking impute operator.
Applying the heuristics from~\Cref{sec:quality} and~\Cref{sec:query-runtime}, we
have
\begin{equation*}
\begin{aligned}
    P(q) &= \sum_{c \in C} \frac{1+\widehat{Var}(c)}{\textsc{Card}(q)} \\
    T(q) &\approx 0.
\end{aligned}
\end{equation*}

@@

Random hot deck imputation is a non-parametric method in which missing values in a column are
replaced with randomly selected complete values from the same column, preserving the
distribution of the complete values. We use the same $P$ as for mean value imputation, but
we modify $T$ to account for the fact that hot-deck is blocking in our implementation.  
\begin{equation*}
\begin{aligned}
    P(q) &= \sum_{c \in C} \frac{1+\widehat{Var}(c)}{\textsc{Card}(q)} \\
    T(q) &= \textsc{Card}(q)
\end{aligned}
\end{equation*}
}
